---
title: "Homework 4"
author: "Sindupa De Silva"
date: "3/12/2021"
output: html_document
---


Included in this homework is an example dataset with the following variables:
  y: response variable (counts of the number of events)
  x1: continuous predictor variable
  x2: categorical predictor variable with 2 levels: "a" and "b"

```{r}

Data <- read.csv ('Homework 4 Data.csv')
head(Data)

```



1. Fit a Poisson regression model that assumes expected count is an interactive function of variables x1 and x2.

```{r}

# Poisson regression model
Fit <- glm(y ~ x1 * x2, family = poisson, data = Data)
summary(Fit)

```



2. Interpret the effect of variable x1 on the expected count when x2 is fixed at level "b". Verify your interpretation in R.

```{r}

# Extract the coefficients 
betas <- coef(Fit)

# log expected count with variable x1 and x2 fixed at level "b"
c_x1 <- betas[1] + betas[2] + betas[3] + betas[4]

# Multiply the x1 effect by 1-unit change
c_x1A <- betas[1] + betas[2] * 1 + betas[3] + betas[4] * 1

c_x1B <- betas[1] + betas[2] * 2 + betas[3] + betas[4] * 2

# For 1-unit change in x1
c_x1B - c_x1A ; log(exp(c_x1B)/ exp(c_x1A))

# Assess proportional change in expected count
exp(c_x1B)/ exp(c_x1A)

# This can be interpretted as:
((exp(c_x1B)/ exp(c_x1A)) - 1) * 100


```
- This shows that there is a 46.1% change in expected count with 1-unit change variable x1, when variable x2 is fixed at level "b".



3. Interpret the effect of variable x2 on the expected count when x1 is fixed at 1. Verify your interpretation in R.

```{r}

# log expected count of variable x2 fixed at level "b"
c_x2B <- betas[1] + betas[2] + betas[3] + betas[4]

c_x2A <- betas[1] + betas[2]

# For 1-unit change in x2
c_x2B - c_x2A ; log(exp(c_x2B)/ exp(c_x2A))

# Assess proportional change
exp(c_x2B)/ exp(c_x2A)

# This can be interpretted as:
((exp(c_x2B)/ exp(c_x2A)) - 1) * 100

```
- This shows that there is a -43.3% change in expected count when variable x2 changes between levels "a" and "b", with variable x1 fixed at 1.



4. Predict the expected count, ± 95% confidence intervals, over the observed range of values of x1, assuming x2 is fixed at level "b".

```{r}

# create vectors of the data
# Vector of x1 values
x1 <- seq(min(Data$x1), max(Data$x1), length.out = 100)

# vector of x2 values fixed at b
x2 <- factor(x = rep('b', times = 100),levels = c('a','b'))

# Combine the data into new data frame
new.data <- data.frame(x1 = x1, x2 = x2)

# Predict values
Pred <- predict.glm(Fit, new.data, type = 'link', se.fit = T)

# Calculate the low limits of our 95% confidence interval
exp(Pred$fit + qnorm(0.025) * Pred$se.fit)

# Calculate the high limits of our 95% confidence interval
exp(Pred$fit + qnorm(0.975) * Pred$se.fit)

# visualizing the predicted values with the high and low confidence intervals

low <- exp(Pred$fit + qnorm(0.025) * Pred$se.fit)
high <- exp(Pred$fit + qnorm(0.975) * Pred$se.fit)

plot(y = exp(Pred$fit), x = new.data$x1, xlab = 'x1',
     ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(low), max(high)), type = 'l')
lines(x = new.data$x1, y = low, lty = 2) 
lines(x = new.data$x1, y = high, lty = 2)
```



5. Predict the expected count, ± 95% confidence intervals, of levels "a" and "b", assuming x1 is fixed at it’s mean.

```{r}

# create vectors of the data
# vector of x2 values of levels "a" and "b"
x2 <- factor(x = Data$x2,levels = c('a','b'))

# Vector of x1 values fixed at its mean
x1 <- seq(min(mean(Data$x1)), max(mean(Data$x1)), length.out = 100)

# Combine the data into new data frame
new.data <- data.frame(x1 = x1, x2 = x2)

# Predict values
Pred <- predict.glm(Fit, new.data, type = 'link', se.fit = T)

# Calculate the low limits of our 95% confidence interval
exp(Pred$fit + qnorm(0.025) * Pred$se.fit)

# Calculate the high limits of our 95% confidence interval
exp(Pred$fit + qnorm(0.975) * Pred$se.fit)

# visualizing the predicted values with the high and low confidence intervals

low <- exp(Pred$fit + qnorm(0.025) * Pred$se.fit)
high <- exp(Pred$fit + qnorm(0.975) * Pred$se.fit)

plot(y = exp(Pred$fit), x = new.data$x2, xlab = 'x2 (a/b)',
     ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(low), max(high)), type = 'l')
lines(x = new.data$x2, y = low, lty = 2) 
lines(x = new.data$x2, y = high, lty = 2)

# I have no idea why this plot is doing this?? Send.Help.

```




