
---
title: "Homework - 2"
author: "Sindupa De Silva"
date: "February 16, 2021"
output: html_document
---


We will use the same beach dataset from last homework. As before, start by fitting a model with OpenBeach as the response, and include an interaction between Year and BeachID.

```{r}
Plover_data <- read.csv("beach.csv")

Fit <- lm (OpenBeach ~ BeachID * Year, data = Plover_data)
summary(Fit)

```

1. Calculate the residuals. Use the summary() function applied to your numeric vector of residuals to
verify that this matches the Residuals: quantiles reported from the summary of your fitted model.
Plot a histogram of your residuals. Do they appear reasonably Gaussian?

$ei = yi ??? yi$
residuals = observed(y) - estimated(y)

```{r}

# Estimated values of y
Est <- fitted.values(Fit)

# Calculated residuals (observed(y) - estimated(y))
resid <- Plover_data$OpenBeach - Est
resid

# Compared with model residuals
residuals(Fit)

# Histogram of residuals
plot(x = resid, y = Plover_data$OpenBeach, type = 'h')

#####  The residuals do not display Gaussian distribution.  #####

```



2. Calculate test statistics for your regression coefficients. Verify by comparing to test statistics reported
from model output.


$t = ??1 ??? B1/ s{??1}$


```{r}

#Calculating test statistics
ts_Fit <- (coef(Fit)) /
  summary(Fit)[['coefficients']][, 'Std. Error']
ts_Fit

#Compared to model output:
summary(Fit)[['coefficients']][, 't value']

```

3. Calculate p-values for your regression coefficients. Verify by comparing to p-values reported from model
output. What are the associated null hypotheses? Do you reject or fail to reject these null hypotheses?
Note: in the lecture slides I reported test statistics are t-distributed with n ??? 2 degrees of freedom.
More generally, these test statistics are t-distributed with n ??? k degrees of freedom, where k is the
number of regression coefficients in your linear model.

```{r}

summary(Fit)
# Degrees of freedom = 56

# For null hypothesis H0: B1 = 0
# Two-tailed
pt(-1 * abs(ts_Fit), df = 56) + (1 - pt(abs(ts_Fit), df = 56))

# Compared to model p-values
summary(Fit)[['coefficients']][,'Pr(>|t|)']

# Intercept and Year p-values are different from model output??

# If the null hypothesis is HA: B1 = 0, with all the observed p-values < 0.05, we reject the null hypothesis.

```

4. Select a single regression coefficient (your choice) and devise a null hypothesis that is different from
the default in lm(). Report the test statistics, your p-value, and whether you reject or fail to reject
your null hypothesis.

```{r}

# Slope coefficient for 'Year'
summary(Fit)[['coefficients']]['Year',]

# New null hypothesis: H0: B1 = 18.37

#Calculating test statistics
ts_Y <- (coef(Fit)[4] - 18.37) /
  summary(Fit)[['coefficients']][ 'Year', 'Std. Error']
ts_Y

# p-value
pt((ts_Y), df = 56)

# The null hypothesis is H0: B1 = 18.37, with the observed p-value < 0.05, we can reject this null hypothesis.

```

5. Interpret output of your fitted model. Tell me how beach area does (or does not change) through time
at each of the 3 beaches.

```{r}

summary(Fit)
betas <- coef(Fit)

#Beach area change for Beach_A
betas[4]

#Beach area change for Beach_B
betas[4] + betas[5]

#Beach area change for Beach_C
betas[4] + betas[6]

```

Looking at the output of the full model, with all p-values < 0.05, we can reject the null hypothesis that H0: B1 = 0 and accept the alternate hypothesis.
- The null hypothesis is that there is no change in beach area per year (H0: B1 = 0)
- The alternate hypothesis is that there is change in beach area per year (HA: B1 =/= 0)

We can see this change in the calculated rate of beach area change per year:
Beach area at Beach "A" decreased by 18.36 ha per year, Beach "B" decreased by 1.32 ha per year, and Beach "C" decreased by 3.99 ha per year.

